---
title: Bailey Simon 2
---

```{r}
library(tidyverse)
```

We'll be looking at Bailey and Simon's first appendix. Here, we assume a universe of risks with three different classes, which we'll call A, B and C. Their populations are 100k, 100k and 50k, respectively with probability of a claim being .05, .1 and .2. If we assume that a claim costs 1,000, then the global risk premium is:

```{r }
claim_cost <- 1e3
class_size <- c(100e3, 100e3, 50e3)
lambda <- c(.05, .1, .2)
mean_lambda <- sum(lambda * class_size) / sum(class_size)
global_premium = claim_cost * mean_lambda
class_premium <- claim_cost * lambda
```

So, we charge everybody 100, which means that class B is getting a fair deal, but no one else is. If we knew what class they were in, we could charge them appropriately. What are the chances that any risk will be in a particular class? That's just the relative frequency.

```{r}
class_probs <- class_size / sum(class_size)
```

```{r}
sum(class_probs * class_premium)
```

So, let's say that we've got someone who has had an accident. What should we charge them? It should be 1,000 times whatever we think their claim frequency is.

```{r}
posterior_class_prob <- (lambda * class_probs) / mean_lambda

posterior_prem <- claim_cost * sum(posterior_class_prob * lambda)

cred <- (posterior_prem - global_premium) / (claim_cost - global_premium)
```

$$
